\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-05-24 manuscript/manuscript.tex
put heuristics into a subsection with further edits

%&latex
\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL
\usepackage{comment}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where
  $C_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is the estimated
  bias term.
\end{enumerate}

@@

In summary, the procedure of the NPBB test is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying circular block bootstrap
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where
  $C_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] -
  F(x; \theta_0^*))$ is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

@@

\section{Appendix}\label{sec:appendix}
In our original setting, we used a block size of $l = \lceil n^{1/3} \rceil$.
However, the optimal block size may vary according to the dependence structure.
To address this, we considered  using the automatic block-length selection
scheme
for
circular block boostrap as described in Section 3.3 of
\citet{politis2004automatic}. Unfortunately, using this setting results in
the selection of
block
sizes that can be quite large (sometimes as large or larger than $n$), leading
to less variation between the bootstrapped
samples. Figures~\ref{fig:large_block_normal} and \ref{fig:large_block_gamma}
demonstrate that for highly temporally
dependent samples (especially when the autocorrelation is negative),
the block size is frequently larger than $\sqrt{n}$.

@@

\bibliographystyle{agsm}

\bibliography{citations}
\end{document}

>>>>

%&latex
\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL
\usepackage{comment}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

\allowdisplaybreaks

\usepackage[pagewise]{lineno}
\linenumbers*[1]
% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
        \expandafter\let\csname old#1\expandafter\endcsname\csname 
        #1\endcsname
        \expandafter\let\csname oldend#1\expandafter\endcsname\csname 
        end#1\endcsname
        \renewenvironment{#1}%
        {\linenomath\csname old#1\endcsname}%
        {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
        \patchAmsMathEnvironmentForLineno{#1}%
        \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
        \patchBothAmsMathEnvironmentsForLineno{equation}%
        \patchBothAmsMathEnvironmentsForLineno{align}%
        \patchBothAmsMathEnvironmentsForLineno{flalign}%
        \patchBothAmsMathEnvironmentsForLineno{alignat}%
        \patchBothAmsMathEnvironmentsForLineno{gather}%
        \patchBothAmsMathEnvironmentsForLineno{multline}%
}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where
  \[
    C_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))
  \]
  is the estimated bias term.
\end{enumerate}

@@

In summary, the procedure of the NPBB test is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying circular block bootstrap
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - K_n(x) \Big\vert.
  \]
  where
  \[
    K_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] -
    F(x; \theta_0^*))
  \]
  is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

\subsection{Heuristic justification for the bias correction term in NPBB}

A clear understanding of the bias correction term $K_n(x)$ is 
essential for appreciating how the NPBB procedure effectively replicates 
the null distribution of the KS statistic in the presence of serial 
dependence and estimated parameters. When the parameters~$\theta$ of a
hypothesized distribution $F(x;\theta)$ are estimated from
the data with $\hat\theta_n$, the null distribution of the KS statistic 
$T_n = \sup_x |\sqrt{n}(F_n(x) - F(x;\hat\theta_n))|$ is altered from its 
standard form. The NPBB procedure aims to replicate this complex null 
distribution.

In the ``bootstrap world,'' each resample yields an empirical distribution 
$F_n^{(b)}$ and a parameter estimate $\hat\theta^{(b)}_n$. The term 
$\sqrt{n}(F_n^{(b)}(x) - F(x;\hat\theta^{(b)}_n))$ is the bootstrap analog of 
the KS process. However, this bootstrap process inherently possesses its own 
mean or systematic deviation. The bias correction $K_n(x)$
estimates this mean.
Specifically, $\mathbb{E}^*[F_n^{(b)}(x)]$ is the average empirical
GXSdistribution function over all bootstrap samples, and
$\theta^*_0 = \mathbb{E}^*[\hat\theta^{(b)}_n]$ is the average of 
the parameter estimates from these bootstrap samples. Thus,
$K_n(x)$ quantifies 
the expected difference, in the bootstrap world, between the empirical
distribution function and the hypothesized distribution function
(evaluated at the average bootstrap parameter estimate).
Subtracting $K_n(x)$ from
$\sqrt{n}(F_n^{(b)}(x) - F(x;\hat\theta^{(b)}_n))$ centers the 
bootstrap KS process. This ensures that the variability of the resulting 
bootstrap statistics $T_n^{(b)}$ more faithfully reflects the variability of 
the original $T_n$ statistic around its true but complex null
distribution.

This approach extends the logic of \citet{babu2004goodness}, who
introduced a similar correction for independent data, to the context
of serially dependent time series via block bootstrapping.
The bias correction term $C_n(x)$ used in the NPB method for
independent data could, in principle, also be approximated by a
bootstrap expectation in the same way as $K_n(x)$.
Nonetheless, under independence, the empirical distribution $F_n(x)$
and the parameter estimate $\hat{\theta}_n$ are already reliable, so
direct calculation of $C_n(x)$ is both simpler and sufficient. The use
of bootstrap expectations as in $K_n(x)$ becomes necessary only when
serial dependence distorts the properties of $F_n(x)$, making
block-bootstrapping-based bias estimation essential.

@@

\section{Appendix}\label{sec:appendix}
In our original setting, we used a block size of $l = \lceil n^{1/3} \rceil$.
However, the optimal block size may vary according to the dependence structure.
To address this, we considered  using the automatic block-length selection
scheme
for
circular block bootstrap as described in Section 3.3 of
\citet{politis2004automatic}. Unfortunately, using this setting results in
the selection of
block
sizes that can be quite large (sometimes as large or larger than $n$), leading
to less variation between the bootstrapped
samples. Figures~\ref{fig:large_block_normal} and \ref{fig:large_block_gamma}
demonstrate that for highly temporally
dependent samples (especially when the autocorrelation is negative),
the block size is frequently larger than $\sqrt{n}$.

@@

\bibliographystyle{asa}
\bibliography{citations}
\end{document}
